---
title: "Color Scales"
author: "Ashwin Malshe"
date: "10/3/2023"
format:
  html: 
    theme: none
---

Note the output. Here, I used a new package called `cleanrmd`. This allows us to change the themes in the output HTML. Read more about it here: https://pkg.garrickadenbuie.com/cleanrmd/reference.html


Install it by running the following command in the console.
install.packages('cleanrmd')

`cleanrmd` includes the following no-class CSS themes:

almond
awsm.css
axist
bamboo
bullframe
holiday
kacit
latex.css
markdown-splendor
markdown-retro
markdown-air
markdown-modest
marx
minicss
new.css
no-class
picocss
sakura
sakura-vader
semantic
simplecss
style-sans
style-serif
stylize
superstylin
tacit
vanilla
water
water-dark
writ

Pick one theme and execute the following function. I am using "bamboo".

```{r}
cleanrmd::use_cleanrmd(name = "bamboo")
```


## Installing and/or loading multiple packages simultaneously

Install package `pacman` by running the following command in the console. **Don't run it in any R chunk below!**

`pacman` enables installing and loading multiple packages in one shot. 

install.packages("pacman")


```{r}
#| label = "setup",
#| include = FALSE

pacman::p_load(tidyverse, lubridate, wesanderson, viridis)
theme_set(theme_minimal())
```

In this code, I will focus on a few common uses of `scale_*_*` functions.

### Change the default color palette

For colors, it is better to use `mpg` data set because it has many categorical variables.
Here I am filtering out observations with `2seater` and `minivan` because the color palettes used in `wesanderson` package has usually 5 distinct colors. So it becomes easier for me to demonstrate the use of the scale function with 5 distinct categories of `class`.


```{r}
p1 = mpg %>% 
  filter(!class %in% c("2seater", "minivan")) %>% 
  ggplot(aes(x = cty, y = hwy, color = class)) +
  geom_point(size = 4, alpha = 0.7)

p1
```


We will use `scale_color_manual()` to change the colors manually.

### `wesanderson` package 

This package is based on the colors used in Wes Anderson movies: https://github.com/karthik/wesanderson
The package developer Karthik Ram is one of the most productive and talented R developers.

```{r}
p1 +
  scale_color_manual(values = wes_palette("Darjeeling1"))
```

Play with other color palettes.

### viridis color palette

**`viridis` is a collection of colorblind-friendly color palettes**

Its developer [Bob Rudis](https://rud.is) is a cybersecurity expert, but he is also a fantastic R developer with several packages under his belt. Check out the `viridis` vignette: https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html


```{r}
p1 +
  scale_color_manual(values = viridis(5))
```

Bob has also added scale functions to directly use these colors.

```{r}
p1 +
  scale_color_viridis(discrete=TRUE)
```

Pick different palettes using `option` argument inside `scale_color_viridis()`

*option* - A character string indicating the colormap option to use. Four options are available: "magma" (or "A"), "inferno" (or "B"), "plasma" (or "C"), "viridis" (or "D", the default option) and "cividis" (or "E").

```{r}
p1 +
  scale_color_viridis(option = "A", discrete=TRUE)
```

Let's take one last example with a continuous color scale.

```{r}
p2 = mpg %>% 
  count(class) %>% 
  ggplot(aes(x = reorder(class, -n), y = n, fill = n)) +
  geom_col() # equivalent to geom_bar(stat = "identity")

p2
  
```
```{r}
p2 +
  scale_fill_viridis(direction = -1)
  
```
