---
title: "Exercise 0"
author: "Collin Real (yhi267)"
format: html
output: pdf_document
---

## Import package, access the studentdata from package, and show part of the data.
```{r}
library(LearnBayes)
data(studentdata)
attach(studentdata)
head(studentdata)
``` 
### 1a) Construct a histogram of this variable using the hist command in R. 
```{r}
hist(studentdata$Dvds, 
     main="DVDs Owned by Students - Histogram", 
     xlab="Total DVDs", 
     ylab="Frequency", 
     col="blue", 
     border="black")
```
### 1b) Summarize this variable using the summary command in R. 
```{r}
summary(studentdata$Dvds)
```
### 1c) Use the table command in R to construct a frequency table of the individual values of Dvds that were observed. If one constructs a barplot of these tabled values using the command `barplot(table(Dvds),col='red')` one will see that particular response values are very popular. Is there any explanation for these popular values for the number of DVDs owned? 
```{r}
table(studentdata$Dvds)
```
```{r}
barplot(table(studentdata$Dvds), 
        main = 'Dvds Owned By Students - Barplot', 
        xlab ='DVDs Count', 
        ylab = 'Student Count', 
        col = 'blue', 
        border = 'black')
```
**1c) Explanation:** There are significant spikes in student count at intervals of 5 & 10, indicating that many students in the survey might not have inputted their exact DVD counts, but rather an estimate. 

### Question 2a) Construct parallel boxplots of the heights using the Gender variable.
```{r}
boxplot(Height ~ Gender, 
        data = studentdata, 
        main = "Heights by Gender - Boxplot", 
        xlab = "Gender", 
        ylab = "Height (inches)", 
        col = c("blue", "green"))
```

### Question 2b) If one assigns the boxplot output to a variable `output=boxplot(Height~Gender)` then output is a list that contains statistics used in constructing the boxplots. Print output to see the statistics that are stored. 
```{r}
output = boxplot(Height ~ Gender)
print(output)
```
### Question 2c) On average, how much taller are male students than female students?
```{r}
avg_male_height = mean(Height[Gender == "male"], 
                       na.rm = TRUE)
avg_female_height = mean(Height[Gender == "female"], 
                         na.rm = TRUE)
height_diff = avg_male_height - avg_female_height

print(paste0("Male height: ", 
             round(avg_male_height, 
                   digits = 2), 
             " inches."))
print(paste0("Female height: ", 
             round(avg_female_height, 
                   digits = 2), 
             " inches."))
print(paste0("On average, male students are ", 
             round(height_diff, 
                   digits = 2), 
             " inches taller than females."))
```

### Question 3a) Construct a scatterplot of ToSleep and WakeUp. 
```{r}
plot(ToSleep, WakeUp, 
     xlab = "Sleep Time", 
     ylab = "WakeUp Time", 
     main = "ToSleep and WakeUp - Scatterplot",
     pch = 19, 
     col = "blue")
```
### Question 3b) Find a least-squares fit to these data using the lm command and then place the least-squares fit on the scatterplot using the abline command.
```{r}
# Plot ToSleep and WakeUp on a scatterplot
plot(ToSleep, WakeUp, 
     xlab = "Sleep Time", 
     ylab = "WakeUp Time", 
     main = "ToSleep and WakeUp - Scatterplot",
     pch = 19,
     col = "blue")

# Find a least-squares fit
fit = lm(WakeUp ~ ToSleep)

# Display the summary
summary(fit)

# Add the least-squares fit line to the scatterplot
abline(fit, col = "black", lwd = 2)
```












