---
title: "Case Study – Customer Acquisition & Retention"
author: "Joaquin, Seth, Collin, Leonel"
format: html
---

## Setup and Data Loading

```{r setup, include=FALSE}
# Load required packages
library(SMCRM)
library(randomForest)
library(rpart)
library(caret)
library(xgboost)
library(tidyverse)
library(vip)  # For variable importance
library(pdp)  # For Partial Dependence Plots (PDPs)
library(pROC)  # For ROC curve and AUC
library(nnet)  # For Neural Networks
library(e1071) # For SVM
library(themis) # For SMOTE

# Load data
data(acquisitionRetention)
head(acquisitionRetention)
```

```{r}
# View data structure
str(acquisitionRetention)
summary(acquisitionRetention$Acquired)
table(acquisitionRetention$Acquired)

# Convert 'Acquired' to a factor variable
acquisitionRetention$Acquired <- as.factor(acquisitionRetention$acquisition)
```


# Data Preprocessing and Train/Test Split

```{r}
acquisitionRetention$Acquired <- sample(c(0, 1), size = nrow(acquisitionRetention), replace = TRUE)
acquisitionRetention$Acquired <- factor(acquisitionRetention$Acquired)


# Check for missing values
sum(is.na(acquisitionRetention$Acquired))

# Check the distribution of the Acquired variable
table(acquisitionRetention$Acquired)
```

```{r}
set.seed(123)
trainIndex <- createDataPartition(acquisitionRetention$Acquired, p = 0.7, list = FALSE)
train <- acquisitionRetention[trainIndex, ]
test <- acquisitionRetention[-trainIndex, ]

# Combine train and test data to handle near-zero variance predictors
combined_data <- rbind(train, test)

# Remove near-zero variance predictors
nzv <- nearZeroVar(combined_data, saveMetrics = TRUE)

# Subset both train and test based on the non-zero variance predictors
train <- train[, !nzv$nzv]
test <- test[, !nzv$nzv]

# Optional: Check if Acquired is imbalanced
table(train$Acquired)
# Check for missing values across all columns
sum(is.na(acquisitionRetention))
```


# Logistic Regression with SMOTE



```{r}
library(tidymodels)
library(glmnet)

# Balance the classes
recipe_smote <- recipe(Acquired ~ ., data = train) %>%
  step_smote(Acquired, over_ratio = 0.5) %>%
  step_normalize(all_predictors())

smote_prep <- prep(recipe_smote, training = train)
train_balanced <- bake(smote_prep, new_data = NULL)

# Fit Regularized Logistic Regression
x <- as.matrix(train_balanced[, -1])
y <- train_balanced$Acquired

cv_fit <- cv.glmnet(x, y, family = "binomial", alpha = 1)
best_lambda <- cv_fit$lambda.min

# Predict on test data
x_test <- as.matrix(test[, -1])
lr_probs <- predict(cv_fit, newx = x_test, s = best_lambda, type = "response")
lr_preds <- ifelse(lr_probs > 0.5, 1, 0)

# Evaluate performance
conf_matrix_lr <- confusionMatrix(factor(lr_preds), test$Acquired)
print(conf_matrix_lr)

# ROC Curve and AUC
roc_lr <- roc(test$Acquired, lr_probs)
plot(roc_lr)
auc(roc_lr)
```


# Random Forest with Hyperparameter Tuning





```{r}
# Tune Random Forest model using caret with ntree parameter
rf_tune <- train(
  Acquired ~ ., data = train, method = "rf",
  trControl = trainControl(method = "cv", number = 5),
  tuneGrid = expand.grid(mtry = c(2, 3, 4, 5)),
  ntree = 500  # Increase the number of trees for better stability
)

# Best Random Forest Model
rf_model <- rf_tune$finalModel

# Predict on test data
rf_preds <- predict(rf_model, test)

# Evaluate performance
conf_matrix_rf <- confusionMatrix(rf_preds, test$Acquired)
print(conf_matrix_rf)

# ROC curve and AUC
roc_rf <- roc(test$Acquired, as.numeric(rf_preds))
plot(roc_rf)
auc(roc_rf)

# Variable Importance Plot
vip(rf_model)
```


# Decision Tree



```{r}
set.seed(123)
trainIndex <- createDataPartition(acquisitionRetention$Acquired, p = 0.7, list = FALSE)
train <- acquisitionRetention[trainIndex, ]
test <- acquisitionRetention[-trainIndex, ]

# Combine train and test data to handle near-zero variance predictors
combined_data <- rbind(train, test)

# Remove near-zero variance predictors
nzv <- nearZeroVar(combined_data, saveMetrics = TRUE)

# Subset both train and test based on the non-zero variance predictors
train <- train[, !nzv$nzv]
test <- test[, !nzv$nzv]

# Fit a Decision Tree model with hyperparameter tuning
dt_model <- rpart(Acquired ~ ., data = train, method = "class", 
                  control = rpart.control(cp = 0.01, minsplit = 20, maxdepth = 5))

# Predict on test data (probabilities for ROC curve)
dt_probs <- predict(dt_model, test, type = "prob")[, 2]
dt_preds <- predict(dt_model, test, type = "class")

# Evaluate performance
conf_matrix_dt <- confusionMatrix(dt_preds, test$Acquired)
print(conf_matrix_dt)

# ROC curve and AUC
roc_dt <- roc(test$Acquired, dt_probs)
plot(roc_dt)
auc(roc_dt)

# Plot the tree
rpart.plot::rpart.plot(dt_model)
```


# Gradient Boosting (XGBoost) with Hyperparameter Tuning

```{r}
# Prepare data for xgboost
train_matrix <- xgb.DMatrix(data = as.matrix(train %>% select(-Acquired)), label = as.numeric(train$Acquired) - 1)
test_matrix <- xgb.DMatrix(data = as.matrix(test %>% select(-Acquired)), label = as.numeric(test$Acquired) - 1)

# Tune Gradient Boosting model using caret
gbm_tune <- train(
  Acquired ~ ., data = train, method = "xgbTree",
  trControl = trainControl(method = "cv", number = 5),
  tuneGrid = expand.grid(
    nrounds = c(100),               # Reduced number of boosting rounds
    max_depth = c(3),               # Reduced max depth
    eta = c(0.1),                   # Learning rate
    gamma = c(0),                   # Minimum loss reduction
    colsample_bytree = c(0.7),      # Proportion of columns for each tree
    min_child_weight = c(1),        # Minimum sum of weights in a child
    subsample = c(0.8)              # Subsample ratio
  )
)

# Best XGBoost Model
gbm_model <- gbm_tune$finalModel

# Predict on test data
gbm_probs <- predict(gbm_model, test_matrix)
gbm_preds <- ifelse(gbm_probs > 0.5, 1, 0)

# Evaluate performance
conf_matrix_gbm <- confusionMatrix(factor(gbm_preds), test$Acquired)
print(conf_matrix_gbm)

# ROC curve and AUC
roc_gbm <- roc(test$Acquired, gbm_probs)
plot(roc_gbm)
auc(roc_gbm)
```


# Support Vector Machine (SVM)

```{r}
# Fit an SVM model
svm_model <- svm(Acquired ~ ., data = train, kernel = "linear")

# Predict on test data
svm_preds <- predict(svm_model, test)

# Evaluate performance
conf_matrix_svm <- confusionMatrix(svm_preds, test$Acquired)
print(conf_matrix_svm)

# ROC curve and AUC
roc_svm <- roc(test$Acquired, as.numeric(svm_preds))
plot(roc_svm)
auc(roc_svm)
```


# k-Nearest Neighbors (kNN)

```{r}
# Install the 'class' package if it's not installed
if (!require(class)) {
  install.packages("class")
}

# Load the 'class' package
library(class)

# Ensure Acquired is a factor variable for classification
train$Acquired <- as.factor(train$Acquired)
test$Acquired <- as.factor(test$Acquired)


# Fit a kNN model
knn_preds <- knn(train = as.matrix(train %>% select(-Acquired)), 
                 test = as.matrix(test %>% select(-Acquired)), 
                 cl = train$Acquired, k = 5)

# Evaluate performance
conf_matrix_knn <- confusionMatrix(knn_preds, test$Acquired)
print(conf_matrix_knn)

# ROC curve and AUC
roc_knn <- roc(test$Acquired, as.numeric(knn_preds))
plot(roc_knn)
auc(roc_knn)
```

Model Comparison

```{r}
# Create a data frame to compare model performance
results <- tibble(
  Model = c("Random Forest", "Decision Tree", "Logistic Regression", "Gradient Boosting", "SVM", "kNN"),
  Accuracy = c(
    mean(rf_preds == test$Acquired),
    mean(dt_preds == test$Acquired),
    mean(lr_preds == test$Acquired),
    mean(gbm_preds == test$Acquired),
    mean(svm_preds == test$Acquired),
    mean(knn_preds == test$Acquired)
  ),
  AUC = c(
    auc(roc_rf),
    auc(roc_dt),
    auc(roc_lr),
    auc(roc_gbm),
    auc(roc_svm),
    auc(roc_knn)
  )
)

# Display results
results %>%
  arrange(desc(Accuracy)) %>%
  knitr::kable()
```


7. Partial Dependence Plots (PDPs)
Generate PDPs for the most important variables in the Random Forest model.


```{r pdp}
# Check column names in the train dataset
colnames(train)
```

```{r}
# Generate Partial Dependence Plot for the "customer" variable
pdp_result <- partial(rf_model, 
                      pred.var = "customer", 
                      train = train,  # Specify the training data
                      plot = TRUE)

# Display the plot
pdp_result


# List of important variables for PDP generation
important_vars <- c("customer", "employees", "revenue")

# Loop through the variables and generate PDPs
for (var in important_vars) {
  print(partial(rf_model, 
                pred.var = var, 
                train = train,  # Specify the training data
                plot = TRUE))
}

```
