---
title: "Merging data"
author: "Pallav Routh"
date: "7/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE, warning = FALSE)
```

Now we will try to merge all available information into one master data set. Again we will rely on dplyr

```{r}
library(dplyr)
```

First we will need to ID variable we can merge on. 

There is no specific order which one can follow to merge multiple files. But, I always like to start with the largest file. So lets try to merge order and order items.

What are variables we can merge on?

```{r}
names(p_orderitems)
```

```{r}
names(p_orders)
```

We can use order ID. We can join p_orderitems from the left of orders.

```{r}
merged_data <-
  p_orderitems |> 
  left_join(p_orders,"order_id")
```

```{r}
nrow(merged_data) == nrow(p_orderitems)
```

Next lets look at reviews data. 

```{r}
names(p_reviews)
```

```{r}
names(merged_data)
```

Again we can use order id.

```{r}
merged_data <-
  p_orderitems |> 
  left_join(p_orders,"order_id") |> 
  left_join(p_reviews,"order_id") 
```

```{r}
nrow(merged_data) == nrow(p_orderitems)
```

Why did it increase?

```{r}
p_reviews |> 
  filter(!is.na(order_id)) |> 
  group_by(order_id) |> 
    summarise(n_obs = n_distinct(review_id)) |> 
  ungroup() |> 
  filter(n_obs > 1)
```

Thats why it grew in size

How can we automate this process?

merged_data <-
  p_orderitems |> 
  left_join(p_orders,"order_id") |> 
  left_join(p_reviews,"order_id") 
  customers - customer id
  sellers - seller_id
  products - product_id

```{r}
names(customers)
names(sellers)
names(products)

names(merged_data)
```

```{r}
merged_data <-
  p_orderitems |> 
  left_join(p_orders,"order_id") |> 
  left_join(p_reviews,"order_id") |> 
  left_join(customers,"customer_id") |> 
  left_join(sellers,"seller_id") |> 
  left_join(products,"product_id")
```


```{r}
all_data <- list(p_orderitems,p_orders,p_reviews,customers,sellers,products)
by_args <- c(rep("order_id",2),"customer_id","seller_id","product_id")

merged_data <- reduce2(all_data,by_args,left_join)
```


Names

```{r}
names(merged_data)
```

Now we will apply pre processing on this merged data. This step is optional.

One thing we need to adjust for is the presence of missing values. The other thing we want to do process variable names. 

```{r}
colSums(is.na(merged_data))
```

Suppose we wish to not deal with them. We can remove these entries. Also notice the few naming errors.

```{r}
p_merged_data <-
  merged_data |> 
    mutate(filter_rows = ifelse(is.na(order_delivered_customer_dt) |
                                is.na(order_approved_at_dt) |
                                is.na(product_category_name),1,0)) |> 
    filter(filter_rows == 0) 

p_merged_data <-
  merged_data |> 
  filter_at(.vars = vars(order_delivered_customer_dt,
                         order_approved_at_dt,
                         product_category_name),
            .vars_predicate = all_vars(!is.na(.)))

p_merged_data <-
    merged_data |> 
    filter_at(.vars = vars(order_delivered_customer_dt,
                         order_approved_at_dt,
                         product_category_name),
             .vars_predicate = all_vars(!is.na(.))) |> 
    select(-review_comm_title,-review_comm_msg) |> 
    rename("product_name_length" = "product_name_lenght",
           "product_description_length" = "product_description_lenght") 

colSums(is.na(p_merged_data))
```

