---
title: "variables"
author: "Pallav Routh"
date: "7/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE, warning = FALSE)
library(dplyr)
```

Creation of variables

1. Order related variables

2. Product related variables

3. Seller specific

4. Customer specific

5. Other factors




1. Order related variables

a. Order delivary time = delivary date - order date

b. Order approval time = order date - approval date

```{r}
estimation_data <- 
  p_merged_data |> 
  mutate(del_time = order_delivered_customer_dt - order_purchase_dt,
         app_time = order_approved_at_dt - order_purchase_dt)
```

2. Product related variables

a. Product dimensions - height, weight and width

b. Product price - the price of the product

c. Product photos - #photos of the product. Here we want to limit the number of cases to under 5 or over 5. 

d. product description length

```{r}
table(p_merged_data$product_photos_qty)
hist(p_merged_data$product_description_length)
my_breaks <- quantile(p_merged_data$product_description_length,probs = c(0:4)/4)

estimation_data <- 
  p_merged_data |> 
  mutate(del_time = order_delivered_customer_dt - order_purchase_dt,
         app_time = order_approved_at_dt - order_purchase_dt,
         photos = ifelse(product_photos_qty > 5,"greater than 5",product_photos_qty),
         des_qntile = cut(product_description_length,
                          breaks = my_breaks,
                          labels = c("25pct","50pct","75pct","100pct"),
                          include.lowest = T)) 

table(estimation_data$des_qntile)
  
```

3. Seller specific

a. Seller business size - the number of products the seller sells

b. Seller competition in a state - the number of sellers within the same state

c. Seller competition in category - the number of sellers within the same product category

```{r}
estimation_data <- 
  p_merged_data |> 
  mutate(del_time = order_delivered_customer_dt - order_purchase_dt,
         app_time = order_approved_at_dt - order_purchase_dt,
         photos = ifelse(product_photos_qty > 5,"greater than 5",product_photos_qty),
         des_qntile = cut(product_description_length,
                          breaks = my_breaks,
                          labels = c("25pct","50pct","75pct","100pct"),
                          include.lowest = T)) |> 
  group_by(seller_id) |> 
    mutate(cust_size = n_distinct(customer_id),
           pdt_size = n_distinct(product_id)) |> 
  ungroup() |> 
  group_by(seller_state) |> 
    mutate(sellersize1 = n_distinct(seller_id)) |> 
  ungroup() |> 
  group_by(product_category_name) |> 
    mutate(sellersize2 = n_distinct(seller_id)) |> 
  ungroup()

estimation_data |> 
  select(product_category_name,sellersize2)
```

4. Customer specific

a. Customer state location - which part of Brazil is the customer located. 
North States: Acre (AC), Amapá (AP), Amazonas (AM), Pará (PA), Rondônia (RO), Roraima (RR), Tocantins (TO). South States: Espírito Santo (ES), Minas Gerais (MG), Rio de Janeiro (RJ), São Paulo (SP), Paraná (PR), Rio Grande do Sul (RS), Santa Catarina (SC)

b. Customers propensity to provide scores in a region - how high do customers in a region usually rate products

```{r}
north <- c("AC","AP","AM","PA","RO","RR","TO")
south <- c("ES","MG","RJ","SP","PR","RS","SC")

estimation_data <- 
  p_merged_data |> 
  mutate(del_time = order_delivered_customer_dt - order_purchase_dt,
         app_time = order_approved_at_dt - order_purchase_dt,
         photos = ifelse(product_photos_qty > 5,"greater than 5",product_photos_qty),
         des_qntile = cut(product_description_length,
                          breaks = my_breaks,
                          labels = c("25pct","50pct","75pct","100pct"),
                          include.lowest = T)) |> 
  group_by(seller_id) |> 
    mutate(cust_size = n_distinct(customer_id),
           pdt_size = n_distinct(product_id)) |> 
  ungroup() |> 
  group_by(seller_state) |> 
    mutate(sellersize1 = n_distinct(seller_id)) |> 
  ungroup() |> 
  group_by(product_category_name) |> 
    mutate(sellersize2 = n_distinct(seller_id)) |> 
  ungroup() |> 
  mutate(stateloc = case_when(customer_state %in% north ~ "north",
                              customer_state %in% south ~ "south",
                              TRUE ~ "other"))

table(estimation_data$stateloc)
```

But we want to save customer state as dummy variables

```{r}
estimation_data |> 
  select(stateloc)  |> 
  mutate(row_no = 1:n(),
         indicator = 1) |> 
  #by pivoting wider we are forcing R to create
  #3 different column - north,couth and others - for every row
  #when south pivots for row 1, it will have 1 but north
  #and other are forced to have NA
  #we will need a unique identifier for every row
  #we need to tell R that this row is different from this row
  tidyr::pivot_wider(names_from = stateloc,
                     values_from = indicator,
                     values_fill = 0)

# if you wanted to create this the traditional way
estimation_data |> 
  select(stateloc)  |> 
  mutate(row_no = 1:n(),
         indicator = ifelse(stateloc == "other",0,1)) |> 
  tidyr::pivot_wider(names_from = stateloc,
                     values_from = indicator,
                     values_fill = 0)

estimation_data <- 
  p_merged_data |> 
  mutate(del_time = order_delivered_customer_dt - order_purchase_dt,
         app_time = order_approved_at_dt - order_purchase_dt,
         photos = ifelse(product_photos_qty > 5,"greater than 5",product_photos_qty),
         des_qntile = cut(product_description_length,
                          breaks = my_breaks,
                          labels = c("25pct","50pct","75pct","100pct"),
                          include.lowest = T)) |> 
  group_by(seller_id) |> 
    mutate(cust_size = n_distinct(customer_id),
           pdt_size = n_distinct(product_id)) |> 
  ungroup() |> 
  group_by(seller_state) |> 
    mutate(sellersize1 = n_distinct(seller_id)) |> 
  ungroup() |> 
  group_by(product_category_name) |> 
    mutate(sellersize2 = n_distinct(seller_id)) |> 
  ungroup() |> 
  mutate(stateloc = case_when(customer_state %in% north ~ "north",
                              customer_state %in% south ~ "south",
                              TRUE ~ "other")) |> 
  mutate(row_no = 1:n(),
         indicator = ifelse(stateloc == "other",0,1)) |> 
  tidyr::pivot_wider(names_from = stateloc,
                     values_from = indicator,
                     values_fill = 0)
```

avg review

```{r}
estimation_data <- 
  p_merged_data |> 
  mutate(del_time = order_delivered_customer_dt - order_purchase_dt,
         app_time = order_approved_at_dt - order_purchase_dt,
         photos = ifelse(product_photos_qty > 5,"greater than 5",product_photos_qty),
         des_qntile = cut(product_description_length,
                          breaks = my_breaks,
                          labels = c("25pct","50pct","75pct","100pct"),
                          include.lowest = T)) |> 
  group_by(seller_id) |> 
    mutate(cust_size = n_distinct(customer_id),
           pdt_size = n_distinct(product_id)) |> 
  ungroup() |> 
  group_by(seller_state) |> 
    mutate(sellersize1 = n_distinct(seller_id)) |> 
  ungroup() |> 
  group_by(product_category_name) |> 
    mutate(sellersize2 = n_distinct(seller_id)) |> 
  ungroup() |> 
  mutate(stateloc = case_when(customer_state %in% north ~ "north",
                              customer_state %in% south ~ "south",
                              TRUE ~ "other")) |> 
  mutate(row_no = 1:n(),
         indicator = ifelse(stateloc == "other",0,1)) |> 
  tidyr::pivot_wider(names_from = stateloc,
                     values_from = indicator,
                     values_fill = 0) |> 
  group_by(customer_state) |> 
    mutate(avg_rev_score = mean(review_score,na.rm = TRUE)) |> 
  ungroup()

estimation_data |> 
  select(customer_state,avg_rev_score)
```


5. Other variables

a. Review scores have some contagion - does review scores for a product on time t-1 have an impact on time t?

```{r}
p_merged_data |> 
  select(seller_id,order_approved_at_dt,product_category_name,review_score) |> 
  arrange(seller_id,order_approved_at_dt)

estimation_data <- 
  p_merged_data |> 
  mutate(del_time = order_delivered_customer_dt - order_purchase_dt,
         app_time = order_approved_at_dt - order_purchase_dt,
         photos = ifelse(product_photos_qty > 5,"greater than 5",product_photos_qty),
         des_qntile = cut(product_description_length,
                          breaks = my_breaks,
                          labels = c("25pct","50pct","75pct","100pct"),
                          include.lowest = T)) |> 
  group_by(seller_id) |> 
    mutate(cust_size = n_distinct(customer_id),
           pdt_size = n_distinct(product_id)) |> 
  ungroup() |> 
  group_by(seller_state) |> 
    mutate(sellersize1 = n_distinct(seller_id)) |> 
  ungroup() |> 
  group_by(product_category_name) |> 
    mutate(sellersize2 = n_distinct(seller_id)) |> 
  ungroup() |> 
  mutate(stateloc = case_when(customer_state %in% north ~ "north",
                              customer_state %in% south ~ "south",
                              TRUE ~ "other")) |> 
  mutate(row_no = 1:n(),
         indicator = ifelse(stateloc == "other",0,1)) |> 
  tidyr::pivot_wider(names_from = stateloc,
                     values_from = indicator,
                     values_fill = 0) |> 
  group_by(customer_state) |> 
    mutate(avg_rev_score = mean(review_score,na.rm = TRUE)) |> 
  ungroup() |> 
  arrange(seller_id,order_approved_at_dt) |> 
  group_by(seller_id,product_category_name) |> 
    mutate(lag_revscore = lag(review_score,1)) |> 
  ungroup()

estimation_data |> 
  select(seller_id,order_approved_at_dt,
         product_category_name,review_score,lag_revscore) |> 
  arrange(seller_id,order_approved_at_dt) 
```


